---
title: "p8105_hw2_mfc2162"
author: "Madeline Cannon"
date: "9/28/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

### Read and clean Mr. Trash Wheel set

```{r}
mr_trash_wheel =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1,
             range = "A2:N406") %>%
              janitor::clean_names() %>%
              drop_na(dumpster) %>%
              mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))

mr_trash_wheel

```


### Read and clean precipitation data

```{r}
precip_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                         sheet = 4, range="A2:B14") %>%
                          janitor::clean_names() %>%
                          mutate(year = 2017) %>%
                          select(month, year, total)

precip_2018 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                         sheet = 3, range="A2:B9") %>%
                          janitor::clean_names() %>%
                          mutate(year = 2018) %>%
                          select(month, year, total)

```


### Combine precipitation datasets and convert month to character variable

```{r}
precip = bind_rows(precip_2017, precip_2018) %>%
                    mutate(month = month.name[month])

precip

```


There are 285 observations in the Mr. Trash Wheel dataset, each corresponding to
a particular dumpster. There are 14 variables, including the date, the weight and
volume of the trash, the frequency of each type of trash (e.g. number of
cigarette butts, glass bottles, grocery bags, etc), and the number of homes
powered. The median number of sports balls in a dumpter in 2017 was
`r median(pull(mr_trash_wheel %>% filter(year == 2017), sports_balls))`.

There are 19 observations in the precipitation dataset, each corresponding to a
particular month. There are three variables: month, year, and total
precipitation. The total precipitation in 2018 was `r sum(pull(precip_2018, total))`.

## Problem 2

### Read and clean pols

```{r}

pols = read_csv("./data/pols-month.csv") %>%
                separate(mon, into = c("year", "month", "day")) %>%
                mutate(month = month.name[as.integer(month)]) %>%
                mutate(president = "gop") %>%
                select(-day)

# Change president to "dem" if prez_dem = 1
for (row in 1:nrow(pols)) {
  if (pols[row, "prez_dem"] == 1) {
    pols[row, "president"] = "dem"
  }
}

# Remove prez_gop and prez_dem
pols = select(pols, -prez_gop, -prez_dem)

pols

```


### Read and clean snp

```{r}

snp = read_csv("./data/snp.csv") %>%
                separate(date, into = c("month", "day", "year")) %>%
                mutate(month = month.name[as.integer(month)]) %>%
                select(year, month, close)

snp

```


### Read and clean unemployment

```{r}

unemployment = read_csv("./data/unemployment.csv") %>%
                        pivot_longer(Jan:Dec,
                                     names_to = "month",
                                     values_to = "unemployment") %>%
                        mutate(month = match(month, month.abb)) %>%
                        mutate(month = month.name[month]) %>%
                        janitor::clean_names() %>%
                        mutate(year = as.character(year))

unemployment

```


### Join the datasets

``` {r}

fivethirtyeight_data = left_join(pols, snp, by = c("year", "month"))

fivethirtyeight_data =
  left_join(fivethirtyeight_data, unemployment, by = c("year", "month"))

fivethirtyeight_data

```


'pols' has 822 observations, each corresponding to a particular date between
1947 and 2015. There are 9 variables that describe whether the president at the
time was a Republican or a Democrat and how many Republican and Democratic
senators, representatives, and governors there were.
'snp' has 787 observations corresponding to dates between 1950 and 2015. There
are 3 variables: year, month, and the S&P stock market index.
'unemployment' has 816 observations corresponding to dates between 1948 and
2015. There are 3 variables: year, month, and the unemployment rate.
'fivethirtyeight_data' has 822 observations (dates between 1947 and 2015) and 11
variables, and is the result of merging the three datasets.

## Problem 3

### Read and clean baby names data

```{r}

baby_names = read_csv("./data/Popular_Baby_Names.csv") %>%
                        janitor::clean_names() %>%
                        mutate(childs_first_name =
                                 str_to_title(childs_first_name))

# Make names of ethnicities consistent
for (row in 1:nrow(baby_names)) {
  if (baby_names[row, "ethnicity"] == "ASIAN AND PACI") {
    baby_names[row, "ethnicity"] = "ASIAN AND PACIFIC ISLANDER"
  }
  if (baby_names[row, "ethnicity"] == "BLACK NON HISP") {
    baby_names[row, "ethnicity"] = "BLACK NON HISPANIC"
  }
  if (baby_names[row, "ethnicity"] == "WHITE NON HISP") {
    baby_names[row, "ethnicity"] = "WHITE NON HISPANIC"
  }
}

# Remove duplicates
baby_names = distinct(baby_names)

baby_names

```


### Olivia table

```{r}

olivia_table = baby_names %>% filter(childs_first_name == "Olivia") %>%
                              select(-childs_first_name, -gender, -count) %>%
                              pivot_wider(
                                names_from = "year_of_birth",
                                values_from = "rank")

olivia_table

```


### Table for most popular male names

```{r}

male_names = baby_names %>% filter(gender == "MALE") %>%
                            filter(rank == 1) %>%
                            select(
                              year_of_birth, ethnicity, childs_first_name) %>%
                            pivot_wider(
                              names_from = "year_of_birth",
                              values_from = "childs_first_name"
                            )

male_names

```


### Scatterplot

```{r}

male_white_2016 = baby_names %>% filter(gender == "MALE") %>%
                                  filter(ethnicity == "WHITE NON HISPANIC") %>%
                                  filter(year_of_birth == 2016)

ggplot(male_white_2016, aes(x = rank, y = count)) + 
  geom_point()

```