---
title: "p8105_hw2_mfc2162"
author: "Madeline Cannon"
date: "9/28/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

### Read and clean Mr. Trash Wheel set

```{r}
mr_trash_wheel =
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1,
             range = "A2:N336") %>%
              janitor::clean_names() %>%
              drop_na(dumpster) %>%
              mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

### Read and clean precipitation data

```{r}
precip_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                         sheet = 4, range="A2:B14") %>%
                          janitor::clean_names() %>%
                          mutate(year = 2017) %>%
                          select(month, year, total)

precip_2018 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                         sheet = 3, range="A2:B9") %>%
                          janitor::clean_names() %>%
                          mutate(year = 2018) %>%
                          select(month, year, total)
```

### Combine precipitation datasets and convert month to character variable

```{r}
precip = bind_rows(precip_2017, precip_2018) %>%
                    mutate(month = month.name[month])
```

There are 285 observations in the Mr. Trash Wheel dataset, each corresponding to
a particular dumpster. There are 14 variables, including the date, the weight and
volume of the trash, the frequency of each type of trash (e.g. number of
cigarette butts, glass bottles, grocery bags, etc), and the number of homes
powered. The median number of sports balls in a dumpter in 2017 was
`r median(pull(mr_trash_wheel %>% filter(year == 2017), sports_balls))`.

There are 19 observations in the precipitation dataset, each corresponding to a
particular month. There are three variables: month, year, and total
precipitation. The total precipitation in 2018 was `r sum(pull(precip_2018, total))`.

## Problem 2

### Read and clean pols-month

```{r}

pols_month = read_csv("./data/pols-month.csv") %>%
              separate(mon, into = c("year", "month", "day")) %>%
              mutate(month = month.name[as.integer(month)]) %>%
              mutate(president = "gop") %>%
              select(-day)

for (row in pols_month) {
  if (pols_month[row, "prez_dem"] == 1) {
    pols_month[row, "president"] = "dem"
  }
}

```